A Hybrid Ensemble Learning Approach to Star-Galaxy Classification

### 2015-05-08 paper/main.tex
udpate paper

\label{firstpage}
\begin{abstract}
%Star-galaxy separation in currently ongoing and future
%large photometric surveys, such as DES and LSST,
%will detect a vast number of unresolved galaxies at faint magnitudes.
There exist a variety of star-galaxy classification techniques,
each with their own strengths and weaknesses.
In this paper, we present a novel meta-classification
framework that combines and fully exploits different techniques
to produce a more robust star-galaxy classification.
To demonstrate this hybrid, ensemble approach,
we combine a purely morphological classifier,
a supervised machine learning method based on random forest,
an unsupervised machine learning method based on self-organizing maps,
and a Hierarchical Bayesian template fitting method.
%Any of these four methods could
%easily be replaced by a similar method
%or be extended to include other existing techniques.
Using data from the CFHTLenS survey,
we consider different scenarios:
when a high-quality training set is available with spectroscopic labels from
DEEP2, SDSS, VIPERS, and VVDS, and
when the demographics of sources in a low-quality training set
do not match the demographics of objects in the test data set.
We demonstrate that our Bayesian combination technique improves
the overall performance over any individual classification method
in these scenarios.
Thus, strategies that combine the predictions of different classifiers
may prove to be optimal in currently ongoing and forthcoming
photometric surveys,
such as the Dark Energy Survey and the Large Synoptic Survey Telescope.

@@

In this paper,
we present a novel star-galaxy classification framework
that combines and fully exploits different classification techniques
to produce a more robust classification.
In particular,
we show that the combination of a morphological separation method,
a template fitting technique, a supervised machine learning method,
and an unsupervised machine learning algorithm
can improve the overall performance over any individual method.
In Section~\ref{section:classification_methods},
we describe each of the star-galaxy classification methods.
In Section~\ref{section:classification_combination_methods},
we describe different classification combination techniques
In Section~\ref{section:data},
we describe the Canada-France Hawaii Telescope Lensing Survey (CFHTLenS)
data set with which we test the algorithms.
In Section~\ref{section:results_and_discussion},
we compare the performance of our combination techniques
to the performance of the individual classification techniques.
Finally, we outline our conclusions in Section~\ref{section:conclusions}.

@@

In this section, we present
four distinct star-galaxy classification techniques.
The first method is a morphological separation method,
which uses a hard cut in the half-light radius vs.\ magnitude plane.
The second method is
a supervised machine learning technique named TPC
(Trees for Probabilistic Classification),
which uses prediction trees and a random forest~\citep{carrascokind2013tpz}.
The third method is
an unsupervised machine learning technique named SOMc,
which uses self-organizing maps (SOMs) and a random atlas
to provide a classification~\citep{carrascokind2014somz}.
The fourth method is
a Hierarchical Bayesian template fitting technique
based on the work by \citet{Fadely2012},
which fits SED templates from star and galaxy libraries
to an observed set of measured flux values.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/morph.eps}
  \caption{Half-light radius vs.\ magnitude.}
  \label{fig:morph}
\end{figure}

@@

Bayesian algorithms use Bayesian inference to quantify
the relative probability that each template matches
the input photometry
and determines a probability estimate by computing
the posterior that a source is a star or a galaxy.
In this work, we have modified and parallelized
a publicly available Hierarchical Bayesian (HB) template fitting
algorithm by \cite{Fadely2012}.
In this section, we provide a brief description
of the HB template fitting technique;
for the details of the underlying HB approach,
we refer the reader to \cite{Fadely2012}.

@@

\noindent
Similarly, we also parametrize the overall prior probability,
$\left(S | \bmath{\theta}\right)$ in Equation~\ref{eq:overall_posterior},
as a weight.
These weights correspond to the hyperparameters,
which can be inferred by sampling
the posterior probability distribution in the hyperparameter space.
For the sampling, we use \textsc{emcee}, a Python implementation of the
affine-invariant Markov Chain Monte Carlo (MCMC) ensemble sampler~\citep{Foreman-Mackey2013}.

@@

Building on the work in the field of ensemble learning,
we combine the predictions from
individual star-galaxy classification techniques
using four combination techniques,
including Bayesian Model Combination (BMC)
and a variant of an ensemble learning technique known as stacking.
The main idea behind ensemble learning is to weight
the predictions from individual models
and combine them to obtain a prediction
that outperforms every one of 
them individually~\citep{rokach2010ensemble}.

@@

Given the variety of star-galaxy classification methods
we are using,
we fully expect the relative performance
of the individual techniques to vary across
the parameter space spanned by the data.
For example, it is reasonable to expect 
supervised techniques to outperform other techniques
in areas of parameter space that are well-populated
with training data.
Similarly, we can expect unsupervised approaches
such as SOM or template fitting approaches 
to generally perform better when training data
is either sparse or unavailable.

We therefore adopt a binning strategy similar to 
\cite{carrascokind2014exhausting}.
In this binning strategy,
we allow different classifier combinations 
in different parts of parameter space
by creating two-dimensional SOM representations of
the full nine-dimensional magnitude-color space:
$u$, $g$, $r$, $i$, $z$, $u-g$, $g-r$, $r-i$, and $i-z$.
A SOM representation can be rectangular, hexagonal, or spherical;
here we choose a 10$\times$10 rectangular topology to facilitate 
visualization as shown in Figure~\ref{fig:som_colors}.
For all combination methods,
we use only the OOB (cross-validation) data contained in each cell
to compute the relative weights for the base classifiers.
The weights within individual cells are then applied to the test data
to make the prediction.

@@

\noindent
where $\hat{y}_i$ is the actual truth value
(\eg 0 or 1) of the $i^{\text{th}}$ data, 
and $y_{i}$ is the probability prediction made by the models.
Thus, a model with the minimum MSE is chosen in each bin
and is assigned a weight of one,
and zero for all other models.
However, the chosen model is allowed to vary between different bins.

@@

\begin{equation} \label{eq:p_ensemble_discrete}
  P \left( e | \bmath{D} \right) \propto 
  P \left( e \right) \left(1 - \epsilon \right)^{N - M_s + M_g}
  \left( \epsilon \right)^{M_s + M_g}.
\end{equation}

@@

We use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{heymans2012cfhtlens,erben2013cfhtlens,hildebrandt2012cfhtlens}.
This catalog consists of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degrees
in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

Our goal here is not to obtain the best classifier
performance; for this we would have fine tuned individual base
classifiers and chosen sophisticated models best suited to the
particular properties of the CFHTLenS data.
For example, \cite{hildebrandt2012cfhtlens} suggest that
all objects with $i > 23$ in the CFHTLenS dataset
may be classified as galaxies
without significant incompleteness and contamination in the galaxy sample.
Although this approach works because
the high Galactic latitude fields of the CFHTLS contain
relatively few stars,
it is very unlikely that such an approach
will meet the science requirements for the quality of
star-galaxy classification in lower latitude, star-crowded fields.
Rather, our goal for the CFHTLenS dataset
is to demonstrate the usefulness
of combining different classifiers even when
the base classifiers may be poor or trained on partial data.

@@

\noindent
where $N_g$ is the number of true galaxies classified as galaxies
and $M_g$ is the number of true galaxies classified as stars.
We define the galaxy \textit{purity} $p_g$ (also known as precision
or positive predictive value)
as the fraction of the number of true galaxies classified as galaxies
out of the total number of objects classified as galaxies, 
\begin{equation}
p_g = \frac{N_g}{N_g + M_s},
\end{equation}

@@

One of the advantages of a probabilistic classification is
that the threshold can be adjusted to produce
a more complete but less pure sample,
or a less complete but more pure one.
To compare the performance of probabilistic classification techniques
with that of morphological separation,
which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
at a certain purity ($p_g = 0.9484$, $p_s = 0.9666$),
we adjust the threshold of probabilistic classifiers
until the galaxy completeness $c_g$ matches
that of morphological separtion
to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
is computed by adjusting the threshold
until the star completeness of each classifier is equal to
that of morphological separation.

@@

In Figure~\ref{fig:som_colors}, we show
the mean CFHTLenS $i$-band magnitude in the top left panel,
and the fraction of stars in each cell in the top right panel.
The bottom two panels show the mean $u-g$ and  $g-r$ colors in each cell.
These two-dimensional maps clearly show
the ability of the SOM to preserve relationships between sources
when it projects the full nine-dimensional space to the two-dimensional map.
We note that these SOM maps should only be used to provide guidance,
as the SOM mapping is a non-linear representation of all magnitudes and colors.

@@

In Figure~\ref{fig:purity_mag}, 
we compare the star and galaxy purity values
for BMC, TPC, and morphological separation
as functions of $i$-band magnitude for the differential counts.
We use the kernel density estimation \cite[KDE;][]{silverman1986density}
with the Gaussian kernel to smooth the fluctuations in the distribution.
Although morphological separation shows a slightly better performance
in galaxy purity at bright magnitudes $i \la 20$,
BMC outperforms both TPC and morphological separation
at faint magnitudes $i \ga 21$.
As the top panel shows,
the number count distribution peaks at $i \sim 22$,
and BMC therefore outperforms both TPC and morphological separation
for the majority of objects.
It is also clear that BMC outperforms TPC over all magnitudes.
BMC can presumably accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.
The bottom panels shows that
the star purity of morphological separation drops to $p_s < 0.8$
at fainter magnitudes $i > 21$.
This is expected, as our crude morphological separation classifies
every object as a galaxy beyond $i > 21$, and
purity measures the number of true stars classified as stars.
It is again clear that BMC outperforms both TPC and morphological separation
in star purity values over all magnitudes.

In Figure~\ref{fig:purity_mag_integrated}, we show
the overall galaxy and star purity values as functions of magnitude
for the integrated counts.
Although morphological separation performs better than TPC
at bright magnitudes, its purity values decrease as 
the magnitudes become fainter, and
TPC eventually outperforms morphological separation by 1--2\% at
$i > 21$.
BMC clearly outperforms TPC, the benchmark base classifier,
and it maintains the overall galaxy purity of 0.980
up to $i \sim 24.5$.

@@

We also show the star and galaxy purity values as a function of
photometric redshift estimate in Figure~\ref{fig:purity_z}.
Photo-$z$ is estimated with
the \textsc{BPZ} algorithm~\citep{benitez2000bayesian}
and provided with
the CFHTLenS photometric redshift catalogue~\citep{hildebrandt2012cfhtlens}.
The advantage of BMC over TPC and morphological separation is
now more pronounced in Figure~\ref{fig:purity_z}.
Although the morphological separation method outperforms BMC
at bright magnitudes in Figure~\ref{fig:purity_mag},
it is clear that BMC outperforms
both TPC and morphological separation over all redshift range.
We also present in Figure~\ref{fig:purity_g_r}
how the star and galaxy purity values vary
as a function of $g-r$ color.
It is again clear that BMC outperforms
both TPC and morphological separation over all $g-r$ color range.

@@

\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           for when the training data set consists of
           only the sources that are in CFHTLS W1 field,
           have spectroscopic labels available from VVDS,
           and have $i < 22$.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \input{tables/metrics_cut.tex}
  \label{table:metrics_cut}
\end{table*}

We present in Table~\ref{table:metrics_cut}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With a smaller training set, our training based methods, TPC and SOMc,
suffer a significant decrease in performance.
The performance of morphological separation and HB
is essentially unchanged from Table~\ref{table:metrics_all}
as they do not depend on the training data.
Without sufficient training data,
the advantage of combining the predictions of different classifiers
is more obvious.
Even WA, the simplest of combination techniques, outperforms
all individual classification techniques in four metrics,
AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
%Interestingly, although TPC does not perform better
%than morphological separation,
%BoM still chooses TPC as the best model within each SOM cell
%and shows exactly the same performance metrics as TPC.
%Stacking and BMC techniques still outperform
%all individual classification techniques,
Alhtough BoM always chooses TPC as the best model
when there is a high-quality training set,
it now chooses various methods and outperforms all base classifiers.
While the performance of the stacking technique is comparable
to that of BMC in Section~\ref{section:rich_training},
stacking now fails to outperform morphological separation.
BMC shows an impressive performance and 
outperforms all other classification techniques in all six metrics.
Overall, the improvements are small but still significant
since these metrics are averaged over the full test data.

@@

We present the star and galaxy purity values as functions of
$i$-band magnitude for the differential counts
in Figure~\ref{fig:purity_mag_cut}.
The normalized density distribution as a function of magnitude
in the top panel
and the stellar distribution in the second panel
clearly show that the demographics of the training set and 
that of the test set are different.
Since the training set is cut at $i < 22$
the density distribution falls off sharply around $i \sim 22$
and has a higher fraction of stars than the test set.
Compared to Figure~\ref{fig:purity_mag},
the purity of TPC suffers a significant decrease in purity.
However, the purity of BMC does not show such a significant drop
and decreases by only 2--5\%.
As suggested by the weight maps in Figure~\ref{fig:weights_cut},
BMC can accomplish this by shifting the relative weights assigned to
each base classifier in different SOM cells.
As the quality of training set worsens,
BMC assigns less weight to training based methods
and more weight to HB and morphological separation.

@@

We also show the star and galaxy purity values as functions of
photo-$z$ in Figure~\ref{fig:purity_z_cut}
and as functions of $g-r$ in Figure~\ref{fig:purity_g_r_cut}.
Compared to Figure~\ref{fig:purity_z} and \ref{fig:purity_g_r},
the performance of BMC becomes worse in some photo-$z$ and $g-r$ bins.
However, this drop in performance seems to be confined to
only a small number of objects in regions of particular parameter space,
and BMC still outperforms both TPC and morphological separation
for the majority of objects.

@@

Compared to Figure~\ref{fig:p_dist_all},
the difference between the posterior star probability distribution of
TPC and that of BMC is now more pronounced in Figure~\ref{fig:p_dist_cut}.
The $P\left(S\right)$ distribution of BMC for true galaxies
falls off sharply at $P\left(S\right)\approx0.95$,
and does not assign a star probability $P(S) \ga 0.95$ to any true galaxies,
while both morphological separation and TPC techniques
classify some true galaxies as stars with absolute certainty.

\begin{figure}
    \centering
  \includegraphics[width=\columnwidth]{figures/p_dist_cut.pdf}
  \caption{Similar to Figure~\ref{fig:p_dist_all},
           but for the reduced training data set.}
  \label{fig:p_dist_cut}
\end{figure}

@@

The combination techniques that we have demonstrated in this work uses
two training based algorithms as base classifiers.
Ideally, the training data should mirror the entire parameter space
occupied by the data to be classified.
Yet we have seen in Section~\ref{section:poor_training}
that the BMC technique does reliably
extrapolate past the limits of the training data,
even when some base classifiers are trained on a low-quality training data set.
In this section, we further investigate
if and where BMC begins to break down
by imposing various magnitude, photo-$z$, and color cuts
to change the size and composition of the training set.

In Figure~\ref{fig:perform_mag_cut}, we present
a visual comparison between different classification techniques,
when various magnitude cuts are applied in the training data,
and the performance is measured on the same test set
from Section~\ref{section:rich_training} and \ref{section:poor_training}.
It is not surprising that the performance of TPC decreases
as we decrease the size of training set
by imposing more restrictive magnitude cuts,
while the performance of HB and morphological separation
is essentially unchanged.
However, the effect of change in size and composition of the training set
is significantly mitigated by the use of the BMC technique.
BMC outperforms both HB and TPC in all four metrics,
even when training is restricted to $i < 20.0$.
BMC also consistently outperforms morphological separation
until we impose a magnitude cut of $i < 20.0$ on the training data,
beyond which point BMC finally performs worse than morphological separation.
It is remarkable that BMC is able to reliably extrapolate
past the training data to $i \sim 24.5$,
the limiting magnitude of the test set, and outperform HB, TPC,
and morphological separation in all performance metrics,
even the demographics of training set do not accurately sample 
the data to be classified.

@@

In Figure~\ref{fig:perform_g_r_cut},
we decrease the size of training set
by keeping red objects and gradually removing blue objects.
A color cut seems to have a more pronounced effect
on the performance of TPC and BMC,
and BMC performs worse than morphological separation
when the training set is restricted to $g - r > 0.4$.
The performance depends more strongly on the color distribution, because 
a significant fraction of blue objects consists of stars,
while objects with fainter magnitudes and higher redshifts
are mostly galaxies.
We can verify this in Figure~\ref{fig:som_colors},
where the darker (higher stellar fraction) cells in the upper middle region
of the stellar fraction map (top right)
have bright magnitudes $i \la 20$ in the $i$-band magnitude map (top left)
and blue colors $g-r \la 0.5$ in the $g-r$ color map (bottom right).
On the other hand, the darker (fainter magnitude) cells
in the lower left corner of the $i$-band magnitude map
have almost no stars in them and
are represented by bright (low stellar fraction)
cells in the stellar fraction map.
Thus, these results indicate that the performance of training based
methods depends more strongly on the composition of training data
than on the size, and it is necessary to have
a sufficient number of the minority class in the training data set
to ensure optimal performance.

>>>>

\label{firstpage}
\begin{abstract}
%Star-galaxy separation in currently ongoing and future
%large photometric surveys, such as DES and LSST,
%will detect a vast number of unresolved galaxies at faint magnitudes.
There exist a variety of star-galaxy classification techniques,
each with their own strengths and weaknesses.
In this paper, we present a novel meta-classification
framework that combines and fully exploits different techniques
to produce a more robust star-galaxy classification.
To demonstrate this hybrid, ensemble approach,
we combine a purely morphological classifier,
a supervised machine learning method based on random forest,
an unsupervised machine learning method based on self-organizing maps,
and a hierarchical Bayesian template fitting method.
%Any of these four methods could
%easily be replaced by a similar method
%or be extended to include other existing techniques.
Using data from the CFHTLenS survey,
we consider different scenarios:
when a high-quality training set is available with spectroscopic labels from
DEEP2, SDSS, VIPERS, and VVDS, and
when the demographics of sources in a low-quality training set
do not match the demographics of objects in the test data set.
We demonstrate that our Bayesian combination technique improves
the overall performance over any individual classification method
in these scenarios.
Thus, strategies that combine the predictions of different classifiers
may prove to be optimal in currently ongoing and forthcoming
photometric surveys,
such as the Dark Energy Survey and the Large Synoptic Survey Telescope.

@@

In this paper,
we present a novel star-galaxy classification framework
that combines and fully exploits different classification techniques
to produce a more robust classification.
In particular,
we show that the combination of a morphological separation method,
a template fitting technique, a supervised machine learning method,
and an unsupervised machine learning algorithm
can improve the overall performance over any individual method.
In Section~\ref{section:classification_methods},
we describe each of the star-galaxy classification methods.
In Section~\ref{section:classification_combination_methods},
we describe different classification combination techniques.
In Section~\ref{section:data},
we describe the Canada-France Hawaii Telescope Lensing Survey (CFHTLenS)
data set with which we test the algorithms.
In Section~\ref{section:results_and_discussion},
we compare the performance of our combination techniques
to the performance of the individual classification techniques.
Finally, we outline our conclusions in Section~\ref{section:conclusions}.

@@

In this section, we present
four distinct star-galaxy classification techniques.
The first method is a morphological separation method,
which uses a hard cut in the half-light radius vs.\ magnitude plane.
The second method is
a supervised machine learning technique named TPC
(Trees for Probabilistic Classification),
which uses prediction trees and a random forest~\citep{carrascokind2013tpz}.
The third method is
an unsupervised machine learning technique named SOMc,
which uses self-organizing maps (SOMs) and a random atlas
to provide a classification~\citep{carrascokind2014somz}.
The fourth method is
a Hierarchical Bayesian (HB) template fitting technique
based on the work by \citet{Fadely2012},
which fits SED templates from star and galaxy libraries
to an observed set of measured flux values.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/morph.pdf}
  \caption{Half-light radius vs.\ magnitude.}
  \label{fig:morph}
\end{figure}

@@

Bayesian algorithms use Bayesian inference to quantify
the relative probability that each template matches
the input photometry
and determine a probability estimate by computing
the posterior that a source is a star or a galaxy.
In this work, we have modified and parallelized
a publicly available Hierarchical Bayesian (HB) template fitting
algorithm by \cite{Fadely2012}.
In this section, we provide a brief description
of the HB template fitting technique;
for the details of the underlying HB approach,
we refer the reader to \cite{Fadely2012}.

@@

\noindent
Similarly, we also parametrize the overall prior probability,
$\left(S | \bmath{\theta}\right)$, in Equation~\ref{eq:overall_posterior},
as a weight.
These weights correspond to the hyperparameters,
which can be inferred by sampling
the posterior probability distribution in the hyperparameter space.
For the sampling, we use \textsc{emcee}, a Python implementation of the
affine-invariant Markov Chain Monte Carlo (MCMC) ensemble sampler~\citep{Foreman-Mackey2013}.

@@

Building on the work in the field of ensemble learning,
we combine the predictions from
individual star-galaxy classification techniques
using four combination techniques.
The main idea behind ensemble learning is to weight
the predictions from individual models
and combine them to obtain a prediction
that outperforms every one of 
them individually~\citep{rokach2010ensemble}.

@@

Given the variety of star-galaxy classification methods
we are using,
we fully expect the relative performance
of the individual techniques to vary across
the parameter space spanned by the data.
For example, it is reasonable to expect 
supervised techniques to outperform other techniques
in areas of parameter space that are well-populated
with training data.
Similarly, we can expect unsupervised approaches
such as SOM or template fitting approaches 
to generally perform better when a training sample
is either sparse or unavailable.

We therefore adopt a binning strategy similar to 
\cite{carrascokind2014exhausting}.
In this binning strategy,
we allow different classifier combinations 
in different parts of parameter space
by creating two-dimensional SOM representations of
the full nine-dimensional magnitude-color space:
$u$, $g$, $r$, $i$, $z$, $u-g$, $g-r$, $r-i$, and $i-z$.
A SOM representation can be rectangular, hexagonal, or spherical;
here we choose a 10$\times$10 rectangular topology to facilitate 
visualization as shown in Figure~\ref{fig:som_colors}.
For all combination methods,
we use only the OOB (cross-validation) data contained in each cell
to compute the relative weights for the base classifiers.
The weights within individual cells are then applied to
the blind test data set to make the prediction.

@@

\noindent
where $\hat{y}_i$ is the actual truth value
(\eg 0 or 1) of the $i^{\text{th}}$ data, 
and $y_{i}$ is the probability prediction made by the models.
Thus, a model with the minimum MSE is chosen in each bin,
and is assigned a weight of one,
and zero for all other models.
However, the chosen model is allowed to vary between different bins.

@@

\begin{equation} \label{eq:p_ensemble_discrete}
  P \left( e | \bmath{D} \right) \propto 
  P \left( e \right) \left(1 - \epsilon \right)^{N - M_s - M_g}
  \left( \epsilon \right)^{M_s + M_g}.
\end{equation}

@@

We use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{heymans2012cfhtlens,erben2013cfhtlens,hildebrandt2012cfhtlens}.
This catalog consists of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degrees
in the four fields (named W1, W2, W3, and W4)
of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

Our goal here is not to obtain the best classifier
performance; for this we would have fine tuned individual base
classifiers and chosen sophisticated models best suited to the
particular properties of the CFHTLenS data.
For example, \cite{hildebrandt2012cfhtlens} suggest that
all objects with $i > 23$ in the CFHTLenS data set
may be classified as galaxies
without significant incompleteness and contamination in the galaxy sample.
Although this approach works because
the high Galactic latitude fields of the CFHTLS contain
relatively few stars,
it is very unlikely that such an approach
will meet the science requirements for the quality of
star-galaxy classification in lower-latitude, star-crowded fields.
Rather, our goal for the CFHTLenS data set
is to demonstrate the usefulness
of combining different classifiers even when
the base classifiers may be poor or trained on partial data.

@@

\noindent
where $N_g$ is the number of true galaxies classified as galaxies,
and $M_g$ is the number of true galaxies classified as stars.
We define the galaxy \textit{purity} $p_g$ (also known as precision
or positive predictive value)
as the fraction of the number of true galaxies classified as galaxies
out of the total number of objects classified as galaxies, 
\begin{equation}
p_g = \frac{N_g}{N_g + M_s},
\end{equation}

@@

One of the advantages of a probabilistic classification is
that the threshold can be adjusted to produce
a more complete but less pure sample,
or a less complete but more pure one.
To compare the performance of probabilistic classification techniques
with that of morphological separation,
which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
at a certain purity ($p_g = 0.9597$, $p_s = 0.9666$),
we adjust the threshold of probabilistic classifiers
until the galaxy completeness $c_g$ matches
that of morphological separtion
to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
is computed by adjusting the threshold
until the star completeness of each classifier is equal to
that of morphological separation.

@@

In Figure~\ref{fig:som_colors}, we show in the top left panel
the mean CFHTLenS $i$-band magnitude in each cell,
and in the top right panel the fraction of stars in each cell.
The bottom two panels show the mean $u-g$ and  $g-r$ colors in each cell.
These two-dimensional maps clearly show
the ability of the SOM to preserve relationships between sources
when it projects the full nine-dimensional space to the two-dimensional map.
We note that these SOM maps should only be used to provide guidance,
as the SOM mapping is a non-linear representation of all magnitudes and colors.

@@

In Figure~\ref{fig:purity_mag}, 
we compare the star and galaxy purity values
for BMC, TPC, and morphological separation
as functions of $i$-band magnitude for the differential counts.
We use the kernel density estimation \cite[KDE;][]{silverman1986density}
with the Gaussian kernel to smooth the fluctuations in the distribution.
Although morphological separation shows a slightly better performance
in galaxy purity at bright magnitudes $i \la 20$,
BMC outperforms both TPC and morphological separation
at faint magnitudes $i \ga 21$.
As the top panel shows,
the number count distribution peaks at $i \sim 22$,
and BMC therefore outperforms both TPC and morphological separation
for the majority of objects.
It is also clear that BMC outperforms TPC over all magnitudes.
BMC can presumably accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.
The bottom panel shows that
the star purity of morphological separation drops to $p_s < 0.8$
at fainter magnitudes $i > 21$.
This is expected, as our crude morphological separation classifies
every object as a galaxy beyond $i > 21$, and
purity measures the number of true stars classified as stars.
It is again clear that BMC outperforms both TPC and morphological separation
in star purity values over all magnitudes.

In Figure~\ref{fig:purity_mag_integrated}, we show
the overall galaxy and star purity values as functions of magnitude
for the integrated counts.
Although morphological separation performs better than TPC
at bright magnitudes, its purity values decrease as 
the magnitudes become fainter, and
TPC eventually outperforms morphological separation by 1--2\% at
$i > 21$.
BMC clearly outperforms both TPC and morphological separation,
and it maintains the overall galaxy purity of 0.980
up to $i \sim 24.5$.

@@

We also show the star and galaxy purity values as functions of
photometric redshift estimate in Figure~\ref{fig:purity_z}.
Photo-$z$ is estimated with
the \textsc{BPZ} algorithm~\citep{benitez2000bayesian}
and provided with
the CFHTLenS photometric redshift catalogue~\citep{hildebrandt2012cfhtlens}.
The advantage of BMC over TPC and morphological separation is
now more pronounced in Figure~\ref{fig:purity_z}.
Although the morphological separation method outperforms BMC
at bright magnitudes in Figure~\ref{fig:purity_mag},
it is clear that BMC outperforms
both TPC and morphological separation over all redshifts.
We also present in Figure~\ref{fig:purity_g_r}
how the star and galaxy purity values vary
as a function of $g-r$ color.
It is again clear that BMC outperforms
both TPC and morphological separation over all $g-r$ colors.

@@

\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           when the training data set consists of
           only the sources that are in CFHTLS W1 field,
           have spectroscopic labels available from VVDS,
           and have $i < 22$.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \input{tables/metrics_cut.tex}
  \label{table:metrics_cut}
\end{table*}

We present in Table~\ref{table:metrics_cut}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With a smaller training set, our training based methods, TPC and SOMc,
suffer a significant decrease in performance.
The performance of morphological separation and HB
is essentially unchanged from Table~\ref{table:metrics_all}
as they do not depend on the training data.
Without sufficient training data,
the advantage of combining the predictions of different classifiers
is more obvious.
Even WA, the simplest of combination techniques, outperforms
all individual classification techniques in four metrics,
AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
%Interestingly, although TPC does not perform better
%than morphological separation,
%BoM still chooses TPC as the best model within each SOM cell
%and shows exactly the same performance metrics as TPC.
%Stacking and BMC techniques still outperform
%all individual classification techniques,
Alhtough BoM always chooses TPC as the best model
when we have a high-quality training set,
it now chooses various methods in different bins
and outperforms all base classifiers.
While the performance of the stacking technique is only slightly worse
than that of BMC when we have a high-quality training set,
stacking now fails to outperform morphological separation.
BMC shows an impressive performance and 
outperforms all other classification techniques in all six metrics.
Overall, the improvements are small but still significant
since these metrics are averaged over the full test data.

@@

We present the star and galaxy purity values as functions of
$i$-band magnitude for the differential counts
in Figure~\ref{fig:purity_mag_cut}.
The normalized density distribution as a function of magnitude
in the top panel
and the stellar distribution in the second panel
clearly show that the demographics of the training set and 
that of the test set are different.
Since the training set is cut at $i < 22$,
the density distribution falls off sharply around $i \sim 22$
and has a higher fraction of stars than the test set.
Compared to the purity values in Figure~\ref{fig:purity_mag},
TPC now suffers a significant decrease in star and galaxy purity.
However, the purity of BMC does not show such a significant drop
and decreases by only 2--5\%.
As suggested by the weight maps in Figure~\ref{fig:weights_cut},
BMC can accomplish this by shifting the relative weights assigned to
each base classifier in different SOM cells.
As the quality of training set worsens,
BMC assigns less weight to training based methods
and more weight to HB and morphological separation.

@@

We also show the star and galaxy purity values as functions of
photo-$z$ in Figure~\ref{fig:purity_z_cut}
and as functions of $g-r$ in Figure~\ref{fig:purity_g_r_cut}.
Compared to Figure~\ref{fig:purity_z} and \ref{fig:purity_g_r},
the performance of BMC becomes worse in some photo-$z$ and $g-r$ bins.
However, this drop in performance seems to be confined to
only a small number of objects in particular regions of
the parameter space,
and BMC still outperforms both TPC and morphological separation
for the majority of objects.

@@

Compared to Figure~\ref{fig:p_dist_all},
the difference between the posterior star probability distribution of
TPC and that of BMC is now more pronounced in Figure~\ref{fig:p_dist_cut}.
The $P\left(S\right)$ distribution of BMC for true galaxies
falls off sharply at $P\left(S\right)\approx0.95$,
and BMC does not assign a star probability
$P(S) \ga 0.95$ to any true galaxies,
On the other hand, both TPC and morphological separation 
classify some true galaxies as stars with absolute certainty.

\begin{figure}
    \centering
  \includegraphics[width=\columnwidth]{figures/p_dist_cut.pdf}
  \caption{Similar to Figure~\ref{fig:p_dist_all}
           but for the reduced training data set.}
  \label{fig:p_dist_cut}
\end{figure}

@@

The combination techniques that we have demonstrated so far use
two training based algorithms as base classifiers.
Ideally, the training data should mirror the entire parameter space
occupied by the data to be classified.
Yet we have seen in Section~\ref{section:poor_training}
that the BMC technique does reliably
extrapolate past the limits of the training data,
even when some base classifiers are trained on a low-quality training data set.
In this section, we further investigate
if and where BMC begins to break down
by imposing various magnitude, photo-$z$, and color cuts
to change the size and composition of the training set.

In Figure~\ref{fig:perform_mag_cut}, we present
a visual comparison between different classification techniques,
when various magnitude cuts are applied on the training data,
and the performance is measured on the same test set
from Section~\ref{section:rich_training} and \ref{section:poor_training}.
It is not surprising that the performance of TPC decreases
as we decrease the size of training set
by imposing more restrictive magnitude cuts,
while the performance of HB and morphological separation
is essentially unchanged.
However, the effect of change in size and composition of the training set
is significantly mitigated by the use of the BMC technique.
BMC outperforms both HB and TPC in all four metrics,
even when the training set is restricted to $i < 20.0$.
BMC also consistently outperforms morphological separation
until we impose a magnitude cut of $i < 20.0$ on the training data,
beyond which point BMC finally performs worse than morphological separation.
It is remarkable that BMC is able to reliably extrapolate
past the training data to $i \sim 24.5$,
the limiting magnitude of the test set, and outperform HB, TPC,
and morphological separation in all performance metrics,
even the demographics of training set do not accurately sample 
the data to be classified.

@@

In Figure~\ref{fig:perform_g_r_cut},
we decrease the size of training set
by keeping red objects and gradually removing blue objects.
A color cut seems to have a more pronounced effect
on the performance of TPC and BMC,
which perform worse than morphological separation
when the training set is restricted to $g - r > 0.4$.
The performance depends more strongly on the color distribution, because 
a significant fraction of blue objects consists of stars,
while objects with fainter magnitudes and higher redshifts
are mostly galaxies.
We can verify this in Figure~\ref{fig:som_colors},
where the darker (higher stellar fraction) cells in the upper middle region
of the stellar fraction map (top right panel)
have bright magnitudes $i \la 20$
in the $i$-band magnitude map (top left panel)
and blue colors $g-r \la 0.5$ in the $g-r$ color map (bottom right panel).
On the other hand, the darker (fainter magnitude) cells
in the right-hand side of the $i$-band magnitude map
have almost no stars in them and
are represented by bright (low stellar fraction)
cells in the stellar fraction map.
Thus, these results indicate that the performance of training based
methods depends more strongly on the composition of training data
than on the size, and it is necessary to have
a sufficient number of the minority class
in the training data set to ensure optimal performance.
